---
title: "README"
output:
  html_document:
    keep_md: yes
---

```{r include=F}
set.seed(12345)
library(optimizeAPA)
library(ggplot2)
```

# redNMix

_redNMix_ is an R package for fitting N-mixture models. The package has several unique features, allowing for:

1. grouped/coarse counts data  
2. parallel processing  
3. applications to large populations

The package can be installed from github:

```{r eval=F}
remotes::install_github("mrparker909/redNMix")
```

```{r}
library(redNMix)
```

# Generate Data

There are helper functions for quickly generating N-mixtures data. These are particularily useful for testing and for simulation studies.

## Closed Population

We generate a closed population with 3 sampling sites and 4 sampling occasions, with site abundance parameter lambda=5, and a probability of detection of 0.80.

```{r}
pop1 = redNMix::gen_Nmix_closed(num_sites = 3,
                                num_times = 4,
                                lambda    = 5,
                                pdet      = 0.8)
```

_pop1_ contains two matrices, the first is the true population size for each sampling:

```{r}
pop1$Ni
```

while the second matrix are the observed counts:

```{r}
pop1$nit
```

## Open Population

We generate an open population with 3 sampling sites and 4 sampling occasions, with initial abundance parameter lambda=5, recruitment parameter gamma=1, survival probability omega=0.25, and a probability of detection of 0.80.

```{r}
pop2 = redNMix::gen_Nmix_open(num_sites = 3,
                              num_times = 4,
                              lambda    = 5,
                              gamma     = 1,
                              omega     = 0.25,
                              pdet      = 0.8)
```

_pop2_ contains two matrices, the first is the true population size for each sampling:

```{r}
pop2$Ni
```

while the second matrix are the observed counts:

```{r}
pop2$nit
```

# Model Fitting

There are many options for model fitting, see the documentation in R for more examples (eg: `?redNMix::fit_red_Nmix_open`). 

## Closed Population Example

```{r cache=T}
out <- redNMix::fit_red_Nmix_closed(nit=pop1$nit, red=1, K=20)
```

```{r}
out
```

The optimization algorithm converged after `r out$steps` iterations. The value of the likelihood function at its maximum is `r out$f`. The parameter estimates are shown under `out$x`, and need to be transformed back to their usual units; lambda was log transformed, so we need to exponentiate it; pdet was logit transformed, so we need to inverse logit it.

The lambda estimate is:

```{r}
exp(out$x[1])
```

The probability of detection estimate is:

```{r}
plogis(out$x[2])
```

The estimates are reasonably close to the true parameter values of 5 and 0.8.

If we want to inspect the convergence paths of the parameters, we should add `keepValues=TRUE` when model fitting:

```{r fig.width=8, fig.height=5, cache=T, warning=F}
out2 <- redNMix::fit_red_Nmix_closed(nit=pop1$nit, red=1, K=20, tolerance=10^-2, keepValues=TRUE)
optimizeAPA::plotConvergence(out2)
```

Notice that we also set the tolerance down from the default of 10^-6 to 10^-2. This reduced the number of iterations from `r out$steps` to `r out2$steps`. In this case the parameter estimates don't change much, so the lower tolerance suffices.

# How to Cite

Parker, M.R.P. (2020). redNMix: An R package for N-mixtures models. R package version `r packageVersion("redNMix")`. https://github.com/mrparker909/redNMix



